Data-Driven Modelling \& \\ Forecasting of Chaotic Systems

### 2022-01-15 Gandhi_Project_Report/draft5.tex
Gandhi's edits of Ch.4

\author{B.L. Nortier} 
\date{January, 2022}
\date{}
\maketitle
\tableofcontents

% \begin{abstract}
%   Abstract goes here.
%   \end{abstract}

@@

  \begin{figure}[ht]
    \includegraphics[scale=0.6]{_tentmap_2.eps}
        \centering
        \captionof{figure}{Three double pendulum heads with initial angle differing by 0,025 radians}\label{fig:T2tentmap}
  \end{figure}

  In general, the graph of  $T^k$ contains $2^{k-1}$ tents.
  Given any point $u$ and an open neighborhood $W \subset[0,1]$ of $u$, we can find $k\in\mathbb{Z}$ large enough so that $T^k(W) = [0,1]$ because we can accommodate a tent whose base is contained in $W$. % \ednote{B: This last phrase doesn't make sense to me. M: The number of tents grow with $k$ and hence this is obvious; I explained in one of our meetings. Does it make sense?} 
  This implies the existence of two points $w_1$ and $w_2$ in $W$ so that $d(T^kw_1,T^kw_2)= 1$, and by the triangle inequality  at least one of the inequalities $d(T^ku,T^kw_1)> \delta$ or $d(T^ku,T^kw_2)> \delta$ hold when $\delta\in (0,\frac{1}{2})$.  So $T$ has sensitive dependence on initial conditions.

@@

  Finally, for every open interval $(a,b)$, the graph of $T^k$ intersects the graph of the identity map on $[0,1]$. This follows from the fact established above that there exists a $k\in\mathbb{N}$ such that $T^k(a,b) = [0,1]$.If the intersection point has coordinates of the form $(p,p)$, the point $p$ is then a fixed point of $T^k$ and therefore also a periodic point of $T$. Since we have found a periodic point in the interval $W$,  the set of periodic points are dense in $T$.
  This follows because any fixed point of $T^k$ is also a periodic point, and we have found this in an arbitrary interval.% \ednote{Expand further on reason for density. M: fixed point of $T^k$ is a periodic point, and we have found in an arbitrary interval.}

@@

First we establish the method of Takens delay embedding. 
In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold)\ednote{Should this be further explained?} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts which will be formalised below.
We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) i.e., we observe  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$. The sequence $\{u_n\}$ represents a scalar time-series and intuitively $\theta$ may be thought to represent a probe inserted into a larger system which is itself only measuring/extracting a small part of the greater system state at time $t=n$. 
Consider for example a thermometer erected to measure the ambient temperature in a local village. This measurement function, the thermometer, is capturing only a single aspect, the temperature, of a much grander dynamical system entailing the present weather of the surrounding area. Even more than that though- it is measuring a miniscule part of the global weather system.

@@

\begin{Definition}
  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP}))] \label{Dfn_NDS}\rm
  A nonautonomous dynamical system (NDS) on a space $X$ is simply a dynamical system comprising a family of maps $\{g_n\}_{n \in \mathbb{Z}}$ where $g_n(\ldotp):=g(u_n,\ldotp):X\to{X}$ is continuous. Each $g_n$ arises as a consequence of an input $u_n$ from the input space $U$, a topological space. 
\end{Definition}

@@

\begin{Example} \rm \label{ex_halfux}
  The only solution  $\{x_n\}_{n\in\mathbb{Z}}$ to the driven system  $g(u,x)=\frac{ux}{2}$, where $X=[0,1]$, $U=[0,1]$,  is the zero solution $x_n\equiv0$.
  To see this, consider any $x_n=a\in[0,1]$ where $a\neq{0}$.  Let $\overline{u}\in{U}$ be an non-zero constant sequence , say $u_n=0.5$. 
  The driven system may be rewritten as $x_{n-1}=\frac{2x_n}{u_{n-1}}=4x_n$ and the  iterates of $x_n$ in backward time will increase by a factor of 4 at each timestep. 
  Thus for some $m\leq{n}$,  $1<x_m$ i.e. $x_m\notin{X}$. So ${\{x_n\}}_{n\in\mathbb{Z}}$ is not a solution and it follows that the only possible solution is the zero solution.
\end{Example}
% \ednote{Nicely written!}

@@

The reachable set itself can defined independently from the fact that $g$ does or does not have the ESP/USP. We note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$. \textbf{(Cite)}.

$X_n$ is exactly the set of points $x$ through which some entire solution $\Psi$ passes at time $n$.

This will lead to a result established later, but which is worth taking note of now: \textit{$g$ being a topological contraction is equivalent to the existence of a unique entire solution.}

% \begin{Definition}
%   [\bf Topological Contraction]\label{Dfn_TopContr}\rm
%   A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subset{U}$, $X_n(\overline{u})$ is a singleton subset of $X$.  
% \end{Definition}

\begin{Remark}
  [\bf Proof of Example~\ref{Ex_exp}] \label{rem_proofEx} \rm
  To see this, we show that for every input $\overline{u}=\{u_n\}_{n\in\mathbb{Z}}$, $g(u_n,\cdot)$ is a contraction map on $X$ for every $k\in\mathbb{Z}$. 
  Indeed $g(u_n,x_n)=x^n$ depends only on the state at time $n$ and so for each input $\overline{u}$, $X_n(\overline{u})=\{x^n\}$ is a singleton subset.
  We then easily conclude that $\{x_n\}\equiv0$ is the only solution. 
\end{Remark} \ednote{B: Is this sufficient?}

Thus far it has been demonstrated that a system may have one or more solutions; one may ask if a driven system always has a solution and, if so, whether it satisfies certain properties such as uniqueness. 
Should the driven system be compact, existence follows immediately as shown in the following result.

\begin{Theorem}\label{Thm_CompactExistence}
 Let $g$ be a driven system.  If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
\end{Theorem}
\begin{proof}
  May be found in \cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}
\end{proof}

One may easily construct many systems with trivial solution-sets, such as $g(u,x)=x$ which has only the constant solution $x$ and so for $U=[-1,1]$, the system would have no solution if $|x|>1$. To refine the scenario, we consider only systems with unique solutions.

@@

% \ednote{M: You can omit the theorem, and just say it is a uniform attractor in the nonautonomous dynamical systems literature \cite{Manju_Nonlinearity}, and say the discussion on nonautonomous attractors is beyond the scope of the thesis.}
% Paraphrased.
Additionally, $g$ having the USP is in our context also equivalent to $g$ being a topological contraction \cite{manjunath2021universal}, or alternatively also that $g$ exhibits the Echo State Property (ESP).

In the majority of Reservoir Computing(RC) literature (a machine learning algorithm approach)\ednote{How should I explain/qualify RC better here? It's mentioned without really explaining.}, 
the ESP is used to describe the concept of fading memory or memory-loss  \cite{jaeger2004harnessing, lu2018attractor}. 
If $g$ possesses the ESP (equivalent in our context to the USP), the we are guaranteed that the whole of an input's left-infinite history will exactly determine the current system state; i.e. there is only ever the possibility of a single reachable state at a given instance of time \cite{jaeger2001echo}

Defined formally, the ESP is intrinsically linked to an input (see \cite{Manju_ESP}) Unfortunately, definitions and results concerning the ESP do not have much practical benefit since a relationship between the network's response and the temporo-statistical properties of the input are not directly described. A more detailed presentation may be found in~\cite{jaeger2001echo}.

We do however consider it important to mention that the ESP is also often discussed in terms of a stability property that holds for all inputs of a system~\cite{manjunath2020stability} and, furthermore, plays a key role in the design and training of recurrent neural networks(RNN) within the field of RC \cite{Manju_ESP}.

\section{Next section?}
\ednote{I feel a break is necessary here, but would hear from you what your thoughts are.}

Having already defined the reachable set $X_U$ as the set of all elements of all entire solutions, we pause briefly in order to fix additional notation.
Defining $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ as the left-infinite subsequence of an input up until time $n$, $\overleftarrow{U}$ is then the notation for all these left-infinite sequences in $U$. 
Moreover, $\cev{u}^{n}v:=(\ldots,u_{n-2} ,u_{n-1}, v)$ will symbolise the input up to time $n$ with $v \in U$ being the specific input value at time $n$. 
The introduction of a new input at time $n$ can be described by the mapping $\sigma_v:   \cev{u}^{n} \mapsto \cev{u}^{n}v$. 
The right-shift map, $r\overleftarrow{U}\to\overleftarrow{U}$, of an input sequence is defined $r: (\cdots, u_{-2},u_{-1}) \mapsto(\cdots, u_{-3},u_{-2})$

The goal now assumes the form of the question on our ability to establish a semi-conjugacy as presented below for the driven system $g$. 
%\ednote{M: I have changed it to a semi-conjugacy since conjugacy amounts to asking a infinite-dimensional space $\overleftarrow{U}$ to be embedded in $X_U$ which is not possible.}

@@

The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows: 
If the causal mapping $H:\overleftarrow{U}{\to}{\overleftarrow{X}_U}$ is injective (in addition to being surjective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$. \ednote{Explain why $H_{-1}$ is continuous} and we refer to $H$ as a \emph{causal embedding} 
% \ednote{M : Causal embedding a dynamical system is different from the definition of causal embedding, so please edit/delete this paragraph and the next theorem and also in the remaining part till Section 1.2}. 
% \begin{Theorem}
%  If $g(\ldotp,x)$ is invertible and has the USP then $H$ is a causal embedding. 
% \end{Theorem}
% \begin{proof}
%   See \cite{manjunath2021universal}.
% \end{proof}
When $g$ has the USP, the diagram below (adopted from~\cite{Manju_Nonlinearity}) illustrates the operation of the mappings $h$ and $H$. The mapping $h:\overleftarrow{U}\to{X_U}$ is also considered an observable as mentioned in the introduction of Chapter \ref{ch3}).

\begin{figure}[ht]
  \includegraphics[scale=0.3]{_actionofh_H.eps}
  \centering
  \captionof{figure}{Ladder-like behaviour of $h$ and $H$ to illustrate the causal mapping. }
  \label{fig:actionh_H}
\end{figure}

We wish to explore the possibility of establishing a conjugacy rather than a mere semi-conjugacy and therefore are interested in the restriction of inputs to the left-infinite orbits of a dynamical system.
The left-infinite orbit spaces are significant because a great amount of information about the original system can be gleaned from their topological structure~\cite{Manju_IEEE}. 
The formal topological structure for the so-called inverse-limit space of chaotic dynamical systems is complicated, and further discussion is relegated to~\cite{kennedy2008inverse_limit, ingram2011inverse}.
Broadly, we may conceive of the inverse-limit space of a dynamical system as a self-map on a subset of an infinite-dimensional space where each point in the inverse-limit space corresponds to a backward orbit of the map $T$. 
Following~\cite{manjunath2021universal, ingram2011inverse} we provide a definition:

\begin{Definition}
  [\bf Inverse-Limit Space, Inverse-Limit System]\label{Dfn_InverseL}\rm
 The subset $\widehat{U}_T\subset \overleftarrow{U}$ defined by \[\widehat{U}_T: = \{ (\ldots,u_{-2},u_{-1}): Tu_{i} = u_{i+1}\]is called the inverse-limit space of $(U,T)$.
 The self-map $\widehat{T}$  induced by $T$ on $\widehat{U}_T$  is defined by \[\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))\] and the resulting dynamical system $(\widehat{U_T}, \widehat{T})$ is called the inverse-limit system of $(U,T)$.
\end{Definition}

The inverse-limit space is well-defined since T : U →U is surjective by assumption.

@@

The driving function $g$ cannot be chosen in a frivolous manner as it could possibly complicate our work. Specifically, one must be careful to avoid a choice of $g$ which would  add complexity to the obtained solution.

When a causal embedding $H$ exists for the driven system $g$, one can map an arbitrary input ${u}$ onto the solution space $X$ without additional distortion or information-loss \textbf{(Cite.)}.
Established such an embedding eliminates the question of possible additional complexity in the solution by guaranteeing that, since the systems are conjugate (semi-conjugate, \textbf{(refer)}), $g$ does not add any (some) complexity to the system.

It is, however, a balancing act as it also undesirable to choose a function $g$ that quenches the temporal structure in $u$ by contracting to such a degree that the ability to recover information from the original system is lost completely.
In the above example (\ref{ex_halfux}), the input's temporal variation cannot be related to the reachable set since $X_U$ consists of a single element; little, if not no, information is encoded.
To obtain a suitably complex function $g$, it is thus desired that the reachable set of a driven system  be large enough to relate to the input. 
 We  reguire that the reachable set of a driven system to be such that the inverse-limit space of $U_T$ can be embedded in some finite self-product of the reachable set of $T$ \textbf{(refer or cite)}. To this end, consider the notion of State-Input (SI) Invertibility.  
 \ednote{(M: Expand here. It stops a bit abruptly. B: Does the above reorganisation read better?)}

\begin{Definition}
  [\bf SI-Invertibility]\label{Dfn_SIinv}\rm
  A map $g$ is said to be SI-Invertible if $g$ is invertible for all $x\in{X}$. Alternatively it may be said that if, given $x_n$ and $x_{n-1}$, $u_{n-1}$ can be uniquely determined from $x_n=g(u_n,x_{n-1})$, then $g$ is said to be SI-invertible.
\end{Definition}

SI-invertibility promises that 'enough' information is retained when $g$ is chosen without introducing additional unwanted complexity. \ednote{(M: Expand why this is true. B: Is this satisfactory?)} 
'Enough' here refers to the fact that we may always recover the previous input value $u_n$ (if we know successive states $x_{n-1}, x_n$).
'Without additional complexity' is guaranteed by the invertibility of the function $g$ that guarantees a one-to-one mapping between $u_n, x_{n-1}$ and $x_n$. 
In obtaining an SI-invertible function, therefore, we are guaranteed a system which will not lose so much information about its previous states in the forward-flow of time that one may not make any credible claims as to the original system's behaviour and properties. 
Simultaneously, it does guarantee us that the important information encoded in previous inputs is indeed preserved.

It is worth at this stage taking note of a specific driven system in the form of a discrete state-space model that has acquired some adherence in applications \cite{Manju_IEEE} and especially those pertaining to Echo State Networks and ESP. The function

\begin{equation}  \label{eqn_driving}
  g(u,x) = (1-a)x + a\overline{\tanh}(Au + \alpha Bx)
\end{equation} 
is both SI-invertible and, if $\alpha\beta<1$, also possesses the USP. 
One may easily show that $g$ is SI-invertible by recovering $u_n$ by 
\begin{equation} \label{eqn_SI_RNN}
  u_{n-1} := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
  \end{equation}
  for $x_{n-1}, x_n$ known.

The proof that the map $g$ possesses the USP is rather involved and hence, instead of being replicated here, may be found in \cite[Th.2]{manjunath2013echo }
This specific driving function $g$ is used in our implementation and is discussed more completely in chapter \ref{sect5}

Despite the ease that one may work with a left-infinite history in the realm of theory, it is impossible to obtain such a sequence in any real-life application.  
One does not in practice, fortunately, need the entire left-infinite history of the input thanks to the Uniform Attraction Property (UAP).

\begin{Definition}
  [\bf Uniform Attraction Property]\label{Dfn_UAP}\rm
  A driven system $g$ has the Uniform Attraction Property (UAP) if, regardless of starting position, all trajectories converge to a single trajectory as time flows forward. 
  This trajectory is also the unique solution sequence $x$ to the input sequence $u$ as mentioned above.
\end{Definition}

This definition is stated in a not completely rigourous manner as the more formal definition makes use of processes, a concept which would take some time to establish and detracts from the principal thrust of this project. 
\ednote{B: Is it very obvious to the reader at this point what the general thrust of the paper is?}

>>>>

\author{B.L. Nortier} 
\date{January, 2022}
\date{}
\maketitle

\begin{abstract}
  Collecting measurements from underlying artificial, natural, and physical systems through observables to realise a dynamical model with greater descriptive power is a well-established practice. Physical systems can often be modelled by instance with differential equations derived from the laws of physics, if we knew the laws.  However, we often do not fully understand the many complexities of the more intricate systems we interact with on a daily basis.
  In this project we start with the hypothesis that an discrete-time dynamical systems, particularly those which are chaotic generates the data. The general learning problem is then formulated and a well-known approach to the problem, alongside its practical limitations are considered.
  We then introduce the notion of driven dynamical systems and establish the theoretical existence of dynamical systems possessing certain properties that guarantee them to be topologically conjugate ('dynamically equivalent') to the dynamics of the underlying system from which measurements were taken. The methodology is practically implemented by making use of a recurrent neural network and using deep learning methods to train on the recurrent neural network states to show that we may obtain long-term consistent predictions for simple physical systems such as the double pendulum as well as a few other attractors exhibiting chaotic behaviour. The machine learning algorithms for modelling the double pendulum thus far have not been successful when it exhibits chaos.

  \end{abstract}

  \tableofcontents

@@

  \begin{figure}[ht]
    \includegraphics[scale=0.6]{_tentmap_2.eps}
        \centering
        \captionof{figure}{Graph of the $T^2$ map with bases at $x=0$, $x=\frac{1}{2}$ and $x=1$ respectively.}\label{fig:T2tentmap}
  \end{figure}

  In general, the graph of  $T^k$ contains $2^{k-1}$ tents.
  Given any point $u$ and an open neighborhood $W \subset[0,1]$ of $u$, we can find $k\in\mathbb{Z}$ large enough so that $T^k(W) = [0,1]$ because we can accommodate a tent whose base is contained in $W$.
  This implies the existence of two points $w_1$ and $w_2$ in $W$ so that $d(T^kw_1,T^kw_2)= 1$, and by the triangle inequality  at least one of the inequalities $d(T^ku,T^kw_1)> \delta$ or $d(T^ku,T^kw_2)> \delta$ hold when $\delta\in (0,\frac{1}{2})$.  So $T$ has sensitive dependence on initial conditions.

@@

  Finally, for every open interval $(a,b)$, the graph of $T^k$ intersects the graph of the identity map on $[0,1]$. This follows from the fact established above that there exists a $k\in\mathbb{N}$ such that $T^k(a,b) = [0,1]$.If the intersection point has coordinates of the form $(p,p)$, the point $p$ is then a fixed point of $T^k$ and therefore also a periodic point of $T$. Since we have found a periodic point in the interval $W$,  the set of periodic points are dense in $T$.
  This follows because any fixed point of $T^k$ is also a periodic point, and we have found this in an arbitrary interval.

@@

First we establish the method of Takens delay embedding. 
In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold)\ednote{Should this manifold be further explained?} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts which will be formalised below.
We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) i.e., we observe  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$. The sequence $\{u_n\}$ represents a scalar time-series and intuitively $\theta$ may be thought to represent a probe inserted into a larger system which is itself only measuring/extracting a small part of the greater system state at time $t=n$. 
Consider for example a thermometer erected to measure the ambient temperature in a local village. This measurement function, the thermometer, is capturing only a single aspect, the temperature, of a much grander dynamical system entailing the present weather of the surrounding area. Even more than that though- it is measuring a miniscule part of the global weather system.

@@

\begin{Definition}
  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP})]\label{Dfn_NDS}\rm
  A nonautonomous dynamical system (NDS) on a space $X$ is simply a dynamical system comprising a family of maps $\{g_n\}_{n \in \mathbb{Z}}$ where $g_n(\ldotp):=g(u_n,\ldotp):X\to{X}$ is continuous. Each $g_n$ arises as a consequence of an input $u_n$ from the input space $U$, a topological space. 
\end{Definition}

@@

\begin{Example} \rm \label{ex_halfux}
  The only solution  $\{x_n\}_{n\in\mathbb{Z}}$ to the driven system  $g(u,x)=\frac{ux}{2}$, where $X=[0,1]$, $U=[0,1]$,  is the zero solution $x_n\equiv0$.
  To see this, consider any $x_n=a\in[0,1]$ where $a\neq{0}$.  Let $\overline{u}\in{U}$ be an non-zero constant sequence , say $u_n=0.5$. 
  The driven system may be rewritten as $x_{n-1}=\frac{2x_n}{u_{n-1}}=4x_n$ and the  iterates of $x_n$ in backward time will increase by a factor of 4 at each timestep. 
  Thus for some $m\leq{n}$,  $1<x_m$ i.e. $x_m\notin{X}$. So ${\{x_n\}}_{n\in\mathbb{Z}}$ is not a solution and it follows that the only possible solution is the zero solution.
\end{Example}

@@

The reachable set itself can defined independently from the fact that $g$ does or does not have the USP. We note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$. \textbf{(Cite)}.
The set $X_n(\overline{u})$ is exactly the set of all points $x$ that have some entire solution $\Psi$ passes through it at time $n$. \ednote{B: Changed sentence for better readability - still understandable?}

We can give an alternate formulation of the set $X_n(\overline{u})$:
suppose, for some fixed input 
$\overline{u}$, we define $g_i = g(u_i,\cdot)$ for all $i\in \mathbb{Z}$, then the set of states
\begin{equation} \label{eqn_association}
X_{n,i}(\overline{u}) := g_{n-1} \circ \cdots g_{i+1} \circ g_i(X).
\end{equation}

As the image of a finite composition of continuous maps, the set $X_{n,i}(\overline{u})$ is compact whenever $X$ is compact and is also nonempty.
Further, $X_{n,i}(\overline{u}) \supset X_{n,i-1}(\overline{u})$. Hence $\bigcap_{i<n}
X_{n,i}(\overline{u})$ is a nested intersection of closed nonempty subsets, and
whenever $X$ is compact, the intersection is nonempty.  \ednote{B: Should I cite here?}

\begin{Definition}
  [\bf Topological Contraction]\label{Dfn_TopContr}\rm
  We say $g$ is a topological contraction if $\bigcap_{i<n}X_{n,i}(\overline{u})$ is a singleton subset of $X$ for each $\overline{u}$. 
\end{Definition}

It turns out that $\bigcap_{i<n}X_{n,i}(\overline{u})$ is identically equal to $X_n(\overline{u})$ that we had defined earlier \textbf{(Cite.)}. This will lead to a result (for. e.g., \cite{manjunath2013echo}) that  \textit{$g$ being a topological contraction is equivalent to the existence of a exactly one entire solution.}

\begin{Remark}
  [\bf Proof of Example~\ref{Ex_exp}] \label{rem_proofEx} \rm
  For the driven system,  
  $g(u,x)=x^2$ the evolution does not depend on the input at all. Hence $g(u,x)$ can be written as a single map $f(x)=x^2$ which is a homeomorphism on $[0,1]$. A left-infinite orbit is defined and it converges to $1$ while the right-infinite orbit converges to $0$. Or in other words the iterates of $f^{-1}$ converges to $1$ while the iterates of $f$ converge to $0$ 
\end{Remark}

Thus far it has been demonstrated that a system may have one or more solutions; one may ask if a driven system always has a solution and, if so, whether it satisfies certain properties such as uniqueness. 
Should the driven system be compact, existence follows immediately.

\begin{Theorem}\label{Thm_CompactExistence}
 Let $g$ be a driven system.  If $X$ is compact, then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
\end{Theorem}
\begin{proof}
  May be found in \cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}
\end{proof}

% One may easily construct many systems with trivial solution-sets, such as $g(u,x)=x$ which has only the constant solution $x$ and so for $U=[-1,1]$, the system would have no solution if $|x|>1$. To refine the scenario, we consider only systems with unique solutions.

@@

% Paraphrased. 
%Additionally, $g$ having the USP is in our context also equivalent to $g$ being a topological contraction \cite{manjunath2021universal}, or alternatively also that $g$ exhibits the Echo State Property (ESP).

\ednote{RC introduced a bit abruptly} Reservoir Computing(RC) refers to a machine learning approach in which input is mapped onto a different, but higher dimensional, space through a system called the reservoir after which a readout measuring the reservoir state is trained.
In the majority of Reservoir Computing(RC) literature \textbf{(Cite)}, a notion of forgetting the states of the reservoir asymptotically is often used.  
Concepts like the echo state property(ESP) and  fading memory \cite{boyd1985fading} are two of these.  
If $g$ possesses the ESP (equivalent in our context to the USP \textbf{(Cite)}), we are guaranteed that the whole of an input's left-infinite history will exactly determine the current system state; i.e. there is only ever the possibility of a single reachable state at a given instance of time \cite{jaeger2001echo,Manju_2020}.
% \cite{Manju_ESP}.
Beyond forgetting the past states, the ESP is also often discussed in terms of a stability property that plays a key role in the robustness of recurrent neural networks(RNN).
The concept, termed input-related stability~\cite{manjunath2020stability}, refers the phenomenon of nearby inputs creating responses (solutions) close to one another.

\section{Establishing Conjugacies}

Having already defined the reachable set $X_U$ as the collection of all elements of all entire solutions, we pause briefly in order to fix additional notation.
Defining $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ as the left-infinite subsequence of an input up until time $n$, $\overleftarrow{U}$ is then the notation for all these left-infinite sequences in $U$. 
Moreover, $\cev{u}^{n}v:=(\ldots,u_{n-2} ,u_{n-1}, v)$ will symbolise the input up to time $n$ with $v \in U$ being the specific input value at time $n$. 
The introduction of a new input at time $n$ can be described by the mapping $\sigma_v:   \cev{u}^{n} \mapsto \cev{u}^{n}v$. 
The right-shift map, $r\overleftarrow{U}\to\overleftarrow{U}$, of an input sequence is defined $r: (\cdots, u_{-2},u_{-1}) \mapsto(\cdots, u_{-3},u_{-2})$

The goal now assumes the form of the question on our ability to establish a semi-conjugacy as presented below for the driven system $g$.

@@

The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows: 
If the causal mapping $H:\overleftarrow{U}{\to}{\overleftarrow{X}_U}$ is injective (in addition to being surjective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$. 
The continuity of $H^{-1}$ follows from the fact that $H$ is itself a continuous and surjective mapping of a compact space $\overleftarrow{U}$ in a Hausdorff space.
and we refer to the map $H$ as a \emph{causal embedding}.

When $g$ has the USP, the diagram below (adopted from~\cite{Manju_Nonlinearity}) illustrates the operation of the mappings $h$ and $H$. The mapping $h:\overleftarrow{U}\to{X_U}$ is also considered an observable as mentioned in the introduction of Chapter \ref{ch3}).

\begin{figure}[ht]
  \includegraphics[scale=0.4]{_actionofh_H.eps}
  \centering
  \captionof{figure}{Ladder-like behaviour of $h$ and $H$ to illustrate the causal mapping. (Figure reproduced from~\cite{Manju_Nonlinearity}) }
  \label{fig:actionh_H}
\end{figure}

The function $H$ contains a countably infinite number of coordinate (or factor maps) since its codomain is $\overleftarrow{X}_U$.  
As it is not practically feasible to consider a left-infinite sequence in $X_U$ (in other words an element of  $\overleftarrow{X}_U$) for learning, we wish to restrict the inputs we consider to a subspace of $\overleftarrow{U}$. 
Particularly, we restrict our attention to the  so-called inverse-limit space of a dynamical system and hope to embed it within some finite product of $X$.

%We wish to explore the possibility of establishing a conjugacy rather than a mere semi-conjugacy and therefore are interested in the restriction of inputs to the left-infinite orbits of a dynamical system.

The left-infinite orbit space is significant because a great amount of information about the original system can be gleaned from its topological structure~\cite{ingram2011inverse,Manju_IEEE}. 
The formal topological structure for the inverse-limit space of a chaotic dynamical system is complicated, and further discussion is relegated to~\cite{kennedy2008inverse_limit, ingram2011inverse}.
Broadly, we may conceive of the inverse-limit system of a dynamical system as a subspace of an infinite-dimensional space where each point in the inverse-limit space corresponds to a backward orbit of the map $T$. 
Following~\cite{manjunath2021universal, ingram2011inverse} we provide a definition:

\begin{Definition}
  [\bf Inverse-Limit Space, Inverse-Limit System]\label{Dfn_InverseL}\rm
 The subset $\widehat{U}_T\subset \overleftarrow{U}$ defined by \[\widehat{U}_T: = \{ (\ldots,u_{-2},u_{-1}): Tu_{i} = u_{i+1}\} \] is called the inverse-limit space of $(U,T)$.
 The self-map $\widehat{T}$  induced by $T$ on $\widehat{U}_T$  is defined by \[\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))\] and the resulting dynamical system $(\widehat{U_T}, \widehat{T})$ is called the inverse-limit system of $(U,T)$.
\end{Definition}

The inverse-limit space is well-defined since $T : U \to{U}$ is surjective by assumption.We now get to the concept of embedding the inverse limit-space of the dynamical system in $X \times X$.

\begin{Definition} \rm
	We say a driven system $g$ \emph{causally embeds} a dynamical system $(U,T)$ if it satisfies the two properties: (i) a universal semi-conjugacy exists and  (ii)  $H_2(\cev{u}) := (h(r\cev{u}),h(\cev{u}))$ embeds the inverse-limit space $\widehat{U}_T$ in $X \times X$. \end{Definition}

It is imperative to take note of a subtlety here: we refer to the \emph{map} $H$ as a (causal) embedding, whereas the driven \emph{system} $g$ (causally) embeds another \emph{system} $(U,T)$.

@@

So far it is not clear if the USP alone ensures the ability to causally embedding a dynamical system and we mention that a driving function $g$ cannot just be chosen in a frivolous manner. This could possibly complicate our work. \ednote{Dié lees vreemd}

%The driving function $g$ cannot be chosen in a frivolous manner as it could possibly complicate our work. Specifically, one must be careful to avoid a choice of $g$ which would  add complexity to the obtained solution.

%When a causal embedding $H$ exists for the driven system $g$, one can map an arbitrary input ${u}$ onto the solution space $X$ without additional distortion or information-loss \textbf{(Cite.)}.

%When an embedding is established, the question of possible additional complexity in the solution is removed by guaranteeing that, since the systems are conjugate (semi-conjugate, \textbf{(refer)}), $g$ does not add any (some) complexity to the system.

%To achieve a causal embedding embedding not just a causal mapping, it is undesirable to choose a function $g$ that quenches the temporal structure in $u$ by contracting to such a degree that the ability to recover information from the original system is lost completely.
In the above example(\ref{ex_halfux}), the input's temporal variation could not be related to the reachable set as $X_U$ only consisted of a single element; little, if not no, information is encoded.
To obtain a suitably complex function $g$, it is thus desired that the reachable set of a driven system  be large enough to relate to the input. 
This is true even for embedding  the inverse-limit space of a dynamical system. \textbf{(B: Expand or refer?)}

 To this end, we recall the notion of State-Input (SI) Invertibility.

\begin{Definition}
  [\bf SI-Invertibility]\label{Dfn_SIinv}\rm
  A driven system $g$ is said to be SI-Invertible if $g(*,x): U \to X$ is invertible for all $x\in X$. Alternatively it may be said that if, given $x_n$ and $x_{n-1}$, $u_{n-1}$ can be uniquely determined from $x_n=g(u_n,x_{n-1})$, then $g$ is said to be SI-invertible.
\end{Definition}

SI-invertibility means that one's choice of $g$ will still ensure 'enough' information is retained without introducing unwanted complexity. 
'Enough' here refers to the fact that we may always recover the previous input value $u_n$ (if we know successive states $x_{n-1}, x_n$).

Subsequently we define the relation $Y_T$ induced by $(U,T)$ on $X_U\times{X_U}$ for a driven system $g$ possessing SI-invertibility.  
To describe the  SDD formally, we consider a dynamical system $T: U \to U$ and define a relation on the reachable set $X_U$, i.e. a subset on $X_U \times X_U$  defined by 
$$Y_T:=\{(x_{n-1},x_n): \{x_k\}_{k\in \mathbb{Z}} \mbox{ is a solution for some orbit of } T \mbox{ and } n \in \mathbb{Z}\}.$$

The following theorem establishes the existence of a well-defined map $G_T$ describing the SDD of the system above.

\begin{Theorem}
If we let $G_T:Y_T\to{Y_T}$ be a map defined by the relation $(x_{n-1},x_n)\mapsto(x_n,x_{n-1})$, then $G_T$ is well-defined (and this results holds even in the absence of $g$ possessing over the USP)  
\end{Theorem}
\begin{proof}
  \textbf{See .....................}
\end{proof}

%'Without additional complexity' is guaranteed by the invertibility of the function $g$ that guarantees a one-to-one mapping between $u_n, x_{n-1}$ and $x_n$. In obtaining an SI-invertible function, therefore, we are guaranteed a system which will not lose so much information about its previous states in the forward-flow of time that one may not make any credible claims as to the original system's behaviour and properties. Simultaneously, it does guarantee us that the important information encoded in previous inputs is indeed preserved.

At this stage it is worth taking note of a specific driven system in the form of a discrete state-space model which has acquired some adherence in applications \cite{Manju_IEEE}- especially those pertaining to Echo State Networks and the ESP. The function 
\begin{equation}  \label{eqn_driving}
  g(u,x) = (1-a)x + a\overline{\tanh}(Au + \alpha Bx)
\end{equation} 
is SI-invertible and, if $\alpha B$ has a spectral norm $<1$, also possesses the USP \cite[Th.2]{manjunath2013echo }. 
It is easy to show that $g$ is SI-invertible by recovering $u_n$ in 
\begin{equation} \label{eqn_SI_RNN}
  u_{n-1} := A^{-1}\bigg(\overline{\tanh}^{^{-1}}\frac{1}{a}\Big(x_{n+1}-(1-a)x_n\Big) \bigg) - \alpha B x_n
  \end{equation}
  when $x_{n-1}$ and $x_n$ are known.

This specific driving function $g$ is used in our implementation and is discussed more completely in chapter~\ref{ch5}
